Type-Safe Modular Parsing

### 2016-11-29 paper/sections/RelatedWork.tex
Related work on parser combinators.

\paragraph*{Parser Combinators} Text.

>>>>

\paragraph*{Parser Combinators} Since [Recursive Programming Techniques 1975] firstly introduced parser generators, and after Wadler discussed more details on backtracking in [1985], parser combinators have been more and more popularly developed and used in the research area of parsing. Among them many parsing libraries produces recursive descent parsers by introducing functional, or more specifically, monadic parser combinators with a foundation on [monadic parser combinators]. Related work includes Parsec [], which is frequently used in Haskell for context-sensitive grammars with infinite lookahead. Nevertheless, left-recursion is known as a big issue in recursive descent parsers, in which case programmers using libraries like Parsec have to manually eliminate left-recursion in the grammar, which is cumbersome, but also it distorts the shape of its old grammar, restricting the modularity of parsing to extreme extent.
There are certainly some approaches to afford the limitation of parser combinators, for example, we have conducted our previous experiment in Haskell, by using open recursion, MRM [] and Parsec. To support direct left-recursive grammars, we made use of the state monad in Parsec to ensure that a left-recursive production rule will not be applied successively in parsing. It turned out to be successful, yet introduced complexity in programming, hence we believe that the parsing library is obligated for left-recursion.

Some more recent papers [Packrat 2006, parsing with derivatives 2011, parser combinator for ambiguous left-recursive grammars] proposed a series of novel parsing techniques, moreover, they did put an eye on the issue of left-recursion. As we explained before, we have selected Packrat Parsing for our prototype implementation, as the combinators are convenient to use in Scala, which is a platform where functional laziness and Object Algebras can perfectly coexist. Also, its support for direct left-recursion is already helpful for designing real-world parsers practically, though [packrat parsing can support left recursion] further demonstrated that general left-recursion is supported from the theoretical point of view. We have also mentioned that the components used in our library can have alternatives, for instance, a different set of parser combinators may support ambiguous and left-recursive grammars, or be applicable to a different subset of context-free grammars or more, or even lead us to another level of performance.

Another important issue in previous papers, to the best of our knowledge, is that none of them deal with modular datatypes or abstract syntax trees. \haoyuan{Well, I need to go to check JS and Patrick Bahr's paper.} The modularity of data structures is yet necessary for modular parsing that supports separate compilation and modular type-checking. In our library, Object Algebras have been adopted for extensible ASTs, together with the behaviors on them.
